### 2018-09-04 resource_justification.tex

\documentclass[11pt]{article}
  \usepackage[a4paper,bindingoffset=0cm,%
              left=2cm,right=2cm,top=2cm,bottom=2cm,%
              footskip=0.5cm]{geometry}
  \usepackage[pdftex]{graphicx}
  \usepackage{multicol}

  %\usepackage{psfig}

  \usepackage{siunitx,amsmath,amssymb}

  \usepackage{sectsty}
  \sectionfont{\fontsize{12}{15}\selectfont}
  \subsectionfont{\fontsize{11}{15}\selectfont}

  %\renewcommand*\sfdefault{phv}
  %\renewcommand{\familydefault}{\sfdefault}
  \makeatletter
  \newenvironment{tablehere}
    {\def\@captype{table}}
    {}

  \newenvironment{figurehere}
    {\def\@captype{figure}}
    {}
  \makeatother

  \begin{document}
  %\title{Broadening the impact of astronomical data handling\vspace{-2ex}}
  %\maketitle
  \setcounter{figure}{0}
  \noindent
  {\LARGE \bf From Stars to Baht: Broadening the economic impact of \\
  astronomical data handling techniques in Thailand}

  \vspace{3mm}
  \noindent
  {\LARGE \bf Justification of resources }
  \vspace{-3mm}
  \section{Directly Incurred Costs: 36,444 (all values in GBP)}
  \vspace{-3mm}
  {\bf Networking events:} We will hold four 3-day networking events in Chiang-Rai to be attended by all co-I's, students and representatives from the external partners (up to 30 people). The first is to introduce the teams to the external partners and to start to understand their needs, a second after six months to discuss and finalise broad specifications of what we will deliver. This will be repeated in year two for the new cohort of students (and any potential new partners). Each event is costed at 2851 and covers: hotel and flight costs for participants not based in Chiang-Rai (1986), transport within Chiang-Rai (230), food and drink (634). Total: 11404. {\bf Transport to partners:} We have budgeted 4468 per year for transport to and accomodation wile at the external partners' sites during the course of the project (8936 total). {\bf MOOC and Training:} Based on prior experience, we have budgeted 10,000 (covering outsourced design, filming, editing, and translating, plus web hosting) for the development of the MOOC to provide training beyond MFU students, and 3052 to hold the training workshops for people enrolled on the MOOC. For the latter, we have budgeted for: food and drink (1030), transport (919), consumables (1103). {\bf Outreach:} During the project, we will host four outreach events in publics to demonstrate to the school-age children the benefits of science in general and data science specifically. Total: 3052

  \vspace{-3mm}
  \section{Investigator Effort: XXXX}
  \vspace{-3mm}
  {\bf Thai salaries:} There are eleven Thai co-Is on the grant, for which we request 0.2FTE each. The 0.2FTE is justified on the basis that all co-I's will attend the 3-day networking events (plus 1 day each to assist with organising the events) and the four-day workshop (20 days total). Each Thai co-I is then required to commit the equivalent of 1 day per fortnight to the project (48 days). This corresponds to 0.18FTE, which we round to 0.2FTE to cover any extra time spent on the project. Total: 46,320 {\bf UK FEC:} Following the same justification, we request 0.2FEC for Mullaney (salary, NI, pension, Estates, Indirect costs), of which STFC will cover 80\%. Total STFC contribution: XXXX

  \vspace{-3mm}
  \section{Indirect and Estate costs at Thai Institutions: 9264}
  \vspace{-3mm}
  As stipulated in the guide, this is calculated at a rate of 20\% of the total Thai investigator salaries.

  \vspace{-3mm}
  \section{Travel and subsistence: 11,424}
  \vspace{-3mm}
  {\bf Thai co-Is/researchers:} We request 2000 per person to cover the costs of two return flights from Thailand to the UK for two research students to visit the GOTO collaboration for one week. We also request 100 per person per day for hotels and subsistence, and 100 per person per year for visas. Total: 7200 {\bf UK PI:} We request 1000 for each return flight for Mullaney to attend each networking event. Mullaney's accommodation and subsistence during the event are already included in the direct costs. Mullaney will stay for a further four nights after each event, for which we request accommodation and subsistence costs of 80 per day. Total STFC contribution (at 80\%): 4224.

  \vspace{-3mm}
  \section{Total Request: XXXX}

  \end{document}